PyUltraLight: A Pseudo-Spectral Solver for Ultralight Dark Matter Dynamics

### 2018-05-19 draft.tex
Additional detail regarding Riemann S ellipsoid

% more complex case: 4 authors, 3 institutions, 2 footnotes
\author[a,b,1]{F. Irst,\note{Corresponding author.}}
\author[c]{S. Econd,}
\author[a,2]{T. Hird\note{Also at Some University.}}
\author[a,2]{and Fourth}

@@

\affiliation[a]{One University,\\some-street, Country}
\affiliation[b]{Another University,\\different-address, Country}
\affiliation[c]{A School for Advanced Studies,\\some-location, Country}

% e-mail addresses: one for each author, in the same order as the authors
\emailAdd{first@one.univ}
\emailAdd{second@asas.edu}
\emailAdd{third@one.univ}
\emailAdd{fourth@one.univ}

@@

An extensive study of the tidal disruption of ULDM solitonic cores orbiting a central potential has recently been undertaken in \cite{Du2018}. Here we reproduce just one of their results as a validity check for \PyUltraLight. In particular, it is found in \cite{Du2018} that an initially spherical solitonic core without self-rotation will gradually spin up to form a tidally-locked irrotational Riemann -S ellipsoid when orbiting a host mass. Figure \ref{fig:riemann} shows the internal velocity field of a solitonic satellite after one complete revolution around a host mass, obtained using \PyUltraLight on a $256^3$ simulation grid. In this case the ratio of host to satellite mass was $\approx 55$. It can be seen that the soliton has become elongated along the radial line connecting to the host, indicating that it is tidally locked. Meanwhile the internal velocity field can be seen to be that of an irrotational Riemann-S ellipsoid \cite{Daller2012}.
\begin{figure}
  \includegraphics[width=1.\textwidth,trim=1cm 1cm 0 1cm,clip]{riemann}
  \caption{Solitonic core after one revolution of a central potential, demonstrating the internal velocity field of an irrotational Riemann-S ellipsoid (where the collective motion of the soliton has been subtracted from the velocity field). Deformation of the spherical density profile into that of an ellipsoid can be seen, with extension along the radial direction (red arrow).}
  \label{fig:riemann}
\end{figure}

>>>>

% more complex case: 4 authors, 3 institutions, 2 footnotes
\author[]{Faber Edwards,}
\author[]{Emily Kendall,}
\author[]{Shaun Hotchkiss,}
\author[]{and Richard Easther}

@@

\affiliation[]{Department of Physics, University of Auckland, Private Bag 92019, Auckland, New Zealand}

% e-mail addresses: one for each author, in the same order as the authors
\emailAdd{faberedwards@gmail.com}
\emailAdd{eken000@aucklanduni.ac.nz}
\emailAdd{s.hotchkiss@auckland.ac.nz}
\emailAdd{r.easther@auckland.ac.nz}

@@

An extensive study of the tidal disruption of ULDM solitonic cores orbiting a central potential has recently been undertaken in \cite{Du2018}. Here, we reproduce just one of their results as an additional validity check for \PyUltraLight. In particular, we reproduce the characteristic internal velocity field of a tidally-locked soliton orbiting a central potential.

When the SP system is recast as a system of hydrodynamical equations, the fluid velocity is given by the gradient of the smoothly-varying phase at any point. This implies that the velocity flow of the SP system is irrotational, $\Nabla\times\vec{v}=0$. When a soliton is initialised in a circular orbit around a Newtonian potential, the initially spherical profile is distorted, elongated along the radial direction of the central potential. Meanwhile, the velocity field corresponding to the overall orbital motion of the soliton is superposed with the internal velocity field, combining so as to produce a net flow with vanishing curl. The family of Riemann-S ellipsoids describe uniformly rotating bodies which satisfy the condition of irrotationality \cite{Chandrasekhar1965}, and it is therefore the characteristic internal velocity field of such an ellipsoid which we expect to arise during the simulation. It is found in \cite{Du2018} that an initially spherical solitonic core without self-rotation will gradually spin up to form a tidally-locked irrotational Riemann-S ellipsoid when orbiting a host mass. We are able to reproduce this result using \PyUltraLight, with Figure \ref{fig:riemann} showing the internal velocity field of a solitonic satellite after one complete revolution around a host mass, obtained using \PyUltraLight on a $256^3$ simulation grid. In this case the ratio of host to satellite mass was $\approx 55$. It can be seen that the soliton has become elongated along the radial line connecting to the host, indicating that it is tidally locked. Meanwhile the internal velocity field can be seen to be that of an irrotational Riemann-S ellipsoid \cite{Daller2012}.
\begin{figure}
  \includegraphics[width=1.\textwidth,trim=1cm 1cm 0 1cm,clip]{riemann}
  \caption{Solitonic core after one revolution of a central potential, demonstrating the internal velocity field of an irrotational Riemann-S ellipsoid (where the collective motion of the soliton has been subtracted from the velocity field). Deformation of the spherical density profile into that of an ellipsoid can be seen, with extension along the radial direction (red arrow).}
  \label{fig:riemann}
\end{figure}
